Text embedding models yield high resolution insights into conceptual knowledge from short multiple choice quizzes

### 2023-02-01 paper/main.tex
overhauled discussion section

Teaching, like effective writing and speaking, is fundamentally about
empathy~\citep{AldrEtal22, StojEtal12, MeyeEtal19}. Great teachers consider
students' interests~\citep{SwarEtal12, Clar10}, backgrounds~\citep{MuijReyn10,
Rose76, denBEtal10}, and working memory capacities~\citep{Allo12}, and flexibly
optimize their teaching strategies within those constraints~\citep{GoodEtal07,
AndeEtal21, John02}. In the classroom, empathizing with students also means
maintaining open lines of communication~\citep{WulfWulf10} by fostering an
environment in which all students feel comfortable speaking up if they have an
exciting new idea, or if they are having trouble understanding
something~\citep{TurnBrai15, GarrRasm14}. In-person instruction also often
entails dynamic student-teacher and student-student interactions. These
in-person interactions can provide the instructor with valuable information
about students' understanding of the course material, beyond what they can
glean solely from exams or assignments~\citep{Engl09, vandEtal10c, HallWals02}.
In turn, this can allow the instructor to adapt their teaching approaches
on-the-fly according to students' questions and behaviors. But what does great
teaching look like in asynchronous online courses, when the instructor
typically prepares course lectures and materials without knowing who will
ultimately be learning from them? Can the empathetic side of teaching be
automated and scaled?

The notion of empathy also related to ``theory of mind'' of other
individuals~\citep{GoldWinn12, KansEtal15, Melt11}. Considering others' unique
perspectives, prior experiences, knowledge, goals, etc., can help us to more
effectively interact and communicate~\citep{ShaoEtal18, StepBaer06, Ratk18}.
The knowledge and learning maps we estimate in our study
(Fig.~\ref{fig:knowledge-maps}) hint at one potential form that an automated
``empathetic'' teacher might take. We imagine automated content delivery
systems that adapt lessons on the fly according to continually updated
estimates of what students know and how quickly they are learning different
conceptual content~\citep[e.g., building on ideas such as][and
others]{AndeSkwa86, Half88, Kuma05, WolzEtal88}.

@@

Ultimately, our work suggests a new line of questions regarding the future of
education: which aspects of teaching can be optimized and/or automated? The
social benefits of face-to-face instruction, such as social interactions,
friendships, and emotional support, cannot (and perhaps should not) be fully
replaced by an automated computer-based system. Nor can modern computer systems
experience emotional empathy in the human sense of the word. On the other hand,
perhaps it is possible to separate out the social aspects of classroom
instruction from the purely learning-related aspects. Our study shows that text
embedding models can uncover detailed insights into students' knowledge and how
it changes over time during learning.  We hope that these advances might help
pave the way for new ways of teaching or delivering educational content
that are tailored to individual students' learning needs and goals.

>>>>

We developed a computational framework that uses short multiple choice quizzes
to provide nuanced insights into what learners know and how their knowledge
changes with training. First, we show that our approach can automatically match
up the conceptual content of individual questions with the corresponding
moments in lecture videos where those concepts were presented
(Fig.~\ref{fig:question-correlations}). Next, we demonstrate how we can
estimate moment-by-moment ``knowledge traces'' that reflect how much knowledge
participants have before and after watching each lecture video
(Fig.~\ref{fig:knowledge-timeseries}). We also show that these knowledge
estimates can generalize to held-out questions (Fig.~\ref{fig:predictions}).
Finally, we use our framework to construct visual maps that provide snapshot
estimates of how much participants know about any concept within the scope of
our text embedding model, and how much their knowledge changes with training
(Fig.~\ref{fig:knowledge-maps}).

@@

Beyond specifically predicting what people \textit{know}, the fundamental ideas
we develop here also relate to the notion of ``theory of mind'' of other
individuals~\citep{GoldWinn12, KansEtal15, Melt11}. Considering others' unique
perspectives, prior experiences, knowledge, goals, etc., can help us to more
effectively interact and communicate~\citep{ShaoEtal18, StepBaer06, Ratk18}.
One could imagine future extensions of our work (e.g., analogous to the
knowledge and learning maps shown in Fig.~\ref{fig:knowledge-maps}), that
attempt to characterize how well-aligned different people's knowledge bases or
backgrounds are. In turn, this might be used to model how knowledge flow not
just between teachers and students, but between friends having a conversation,
individuals out on a first date, participants at a business meeting, doctors
and patients, experts and non-experts, political allies and adversaries, and
more. For example, the extent to which two people's knowledge maps ``match'' in
a given region of text embedding space might serve as a predictor of how
effectively the people will communicate about the corresponding conceptual
content.

Ultimately, our work suggests a rich new line of questions about the ``form''
of knowledge, how knowledge changes over time, and how we might map out the
full space of what an individual knows. Our finding that detailed estimates
about knowledge may be obtained from short quizzes shows one way that
traditional approaches to evaluation in education might be extended. We hope
that these advances might help pave the way for new ways of teaching or
delivering educational content that are tailored to individual students'
learning needs and goals.
